CYO Project Report - Brazilian Ecommerce

### 2021-07-21 finalReport.Rmd
intermediate state for record keeping

@@

```{r Explore volume of product group with number_of_hours, echo=FALSE}
# let's plot errorbar for number of hours vs volume of product group
training_dataset %>% mutate(volume_group=round(volume/1000)+1) %>% mutate(volume_group=as.factor(volume_group)) %>% group_by(volume_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=volume_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
```
It does not influencee much the number of hours

@@

@@

@@

  set.seed(1,sample.kind='Rounding')
  test_weight_indices <- createDataPartition(train_subset$number_of_hours,p=0.2,times=5,list=FALSE)

seller_id_effects_list <- lapply(1:5, function(index) {
    train_weight_set <- train_subset[-test_weight_indices[,index],]

@@

sie <- seller_id_effects_list[[1]]
for (index in 2:5) {
   sie <- sie %>% full_join(seller_id_effects_list[[index]],by='seller_id', suffix=c("1",index))
}

@@

train_subset <- train_subset %>% left_join(seller_id_effects,by='seller_id') %>% select(-product_category_name,-seller_id) %>% mutate(seller_id_effect=if_else(is.na(seller_id_effect),global_mean,seller_id_effect))

test_subset <- test_subset %>% left_join(seller_id_effects,by='seller_id') %>% select(-product_category_name,-seller_id)
```

@@

  set.seed(1,sample.kind='Rounding')
  test_weight_indices <- createDataPartition(train_subset$number_of_hours,p=0.2,times=5,list=FALSE)

customer_state_effects_list <- lapply(1:5, function(index) {
    train_weight_set <- train_subset[-test_weight_indices[,index],]

@@

cse <- customer_state_effects_list[[1]]
for (index in 2:5) {
   cse <- cse %>% full_join(customer_state_effects_list[[index]],by='customer_state', suffix=c("1",index))
}

@@

tuningControl=trainControl(method = "cv", number = 3, p = .8)
knn_fit <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(27,34,41)),trControl=tuningControl)

@@

RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
#184.7906 for k=34.

@@

```{r correlation, eval=FALSE}
cor(train_subset %>% select(price,freight_value,customer_loc_effect,customer_state_effect,seller_id_effect,distance_between_two_points,mean_delay_in_handover,volume,number_of_hours,product_weight_g),use="pairwise.complete.obs") %>% knitr::kable()
```

@@

```{r rf model, echo = FALSE}
set.seed(seed=1,sample.kind = 'Rounding')
tuningControl=trainControl(method = "cv", number = 2, p = .8)
rf_fit <- train(number_of_hours~., train_subset, method='rf',tuneGrid = data.frame(mtry = seq(5, 40, 10)),trControl=tuningControl,ntree=250)
#rf_fit <- train(number_of_hours~., train_subset %>% select(-seller_state), method='rf',tuneGrid = data.frame(mtry = seq(5, 50, 10)),trControl=tuningControl,ntree=250) # nodesize = 50,maxnodes=2000,ntree=200

@@

RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
# 175.0988 for mtry 15. Technically there are total of 82 variables due to 73 product categories and 9 other predictors. The seller_state was not of much help. it was actually increasing RMSE by .6. So have ignored it.

@@

RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
# 185.1074 uses GCV.Cp method and select=FALSE

@@

@@

@@

  # cleanup
  if (cleanup) {
    train_df <- train_df %>% select(-product_category_name,-seller_id)
    test_df <- test_df %>% select(-product_category_name,-seller_id)
  }

@@

@@

validation_dataset <- validation_dataset %>%
  semi_join(training_dataset,by='product_category_name') %>% 
  semi_join(training_dataset,by='seller_id') %>% 
  semi_join(training_dataset,by='customer_state') %>%
  semi_join(training_dataset,by='customer_loc')

# cleanup
training_dataset <- training_dataset %>%
  select(-handed_over_to_carrier_date,-shipping_limit_date,-product_category_name,-seller_id,-customer_state,-customer_loc,-customer_zip_code_prefix)
validation_dataset <- validation_dataset %>%
  select(-handed_over_to_carrier_date,-shipping_limit_date,-product_category_name,-seller_id,-customer_state,-customer_loc,-customer_zip_code_prefix)
```

@@

RMSE((validation_dataset %>% pull(number_of_hours) ),predictions)
# 190.184 for mtry=15, ntree=500
```

>>>>

```{r smooth fit for price, echo=FALSE}
training_dataset %>% ggplot(aes(x=price,y=number_of_hours)) + geom_smooth() 
```

@@

```{r Explore volume of product group with number_of_hours, echo=FALSE}
# let's plot errorbar for number of hours vs volume of product group
training_dataset %>% mutate(volume_group=round(volume/5000)+1) %>% mutate(volume_group=as.factor(volume_group)) %>% group_by(volume_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=volume_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
```
It does not influencee much the number of hours

@@

== TODO additive smoothing/Bayes estimator with target encoding

@@

```{r Apply Target encoding for product_category_name, echo = FALSE}
#TODO you may need to remove customer_state and seller_state from dataset train_subset_for_encoding before proceeding further.

  train_subset_for_encoding <- train_subset[1:30000,] %>% select(-seller_id,-customer_loc, -customer_zip_code_prefix, -customer_state)

  test_weight_set <- test_weight_set %>%
    semi_join(train_weight_set,by='product_category_name')

  weight_rmses <- sapply(weights, function(weight) {
      product_category_effects <- train_weight_set %>% 
        group_by(product_category_name) %>% 
        summarise(mean=mean(number_of_hours),n=n()) %>%
        mutate(mean=(n*mean+weight*global_mean)/(n+weight)) %>% 
        select(-n) %>% rename()

      train_weight_set <- train_weight_set %>% left_join(product_category_effects,by='product_category_name') %>% rename(product_category_effect=mean) %>% select(-product_category_name)

      test_weight_set <- test_weight_set %>% left_join(product_category_effects,by='product_category_name') %>% rename(product_category_effect=mean) %>% select(-product_category_name)

      tuningControl=trainControl(method = "cv", number = 3, p = .8)
      knn_fit_for_weight <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(27,34,41)),trControl=tuningControl)

      knn_fit_for_weight
      print(knn_fit_for_weight)

target_encoding_weights <- target_encoding_weights %>% add_row(effect='product_category',weight=weight_to_use)
```

```{r Post regularization for product_category effect, echo=FALSE}

# prepare product_category_effect based on regularisation weight we have identified
  global_mean <- mean(train_subset$number_of_hours)

product_category_effects_list <- lapply(1:10, function(index) {
    train_weight_set <- train_subset[-test_weight_indices[,index],]

    product_category_effects <- train_weight_set %>% 
      group_by(product_category_name) %>%
      summarise(mean=mean(number_of_hours),n=n()) %>%
      mutate(mean=(n*mean+weight_to_use*global_mean)/(n+weight_to_use)) %>% 
      select(-n) %>% rename()

    product_category_effects
  }
)

pce <- product_category_effects_list[[1]]
for (index in 2:10) {
   pce <- pce %>% full_join(product_category_effects_list[[index]],by='product_category_name', suffix=c("1",index))
}

# wherever value is NA, we have used global mean
pce <- pce %>% mutate(across(-product_category_name,~if_else(is.na(.x),global_mean,.x)))

# as different partitions may not have exactly same data set combined the results
product_category_effects <- tibble(pce[,1], product_category_effect=rowMeans(pce[,-1],na.rm = TRUE))

train_subset <- train_subset %>% left_join(product_category_effects,by='product_category_name') %>% select(-product_category_name) %>% mutate(product_category_effect=if_else(is.na(product_category_effect),global_mean,product_category_effect))

test_subset <- test_subset %>% left_join(product_category_effects,by='product_category_name') %>% select(-product_category_name)
```

@@

seller_id_effects_list <- lapply(1:10, function(index) {
    train_weight_set <- train_subset[-test_weight_indices[,index],]

@@

# wherever value is NA, we have used global mean
sie <- sie %>% mutate(across(-seller_id,~if_else(is.na(.x),global_mean,.x)))

@@

train_subset <- train_subset %>% left_join(seller_id_effects,by='seller_id') %>% select(-seller_id) %>% mutate(seller_id_effect=if_else(is.na(seller_id_effect),global_mean,seller_id_effect))

test_subset <- test_subset %>% left_join(seller_id_effects,by='seller_id') %>% select(-seller_id)
```

@@

customer_state_effects_list <- lapply(1:10, function(index) {
    train_weight_set <- train_subset[-test_weight_indices[,index],]

@@

# wherever value is NA, we have used global mean
cse <- cse %>% mutate(across(-customer_state,~if_else(is.na(.x),global_mean,.x)))

@@

tuningControl=trainControl(method = "cv", number = 5, p = .8)
knn_fit <- train(number_of_hours~., scaled_train_subset, method='knn',tuneGrid=data.frame(k=c(56,63,70,77)),trControl=tuningControl)

@@

RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
#183.2447 for k=70

@@

```{r correlation, eval=FALSE}
cor(train_subset %>% select(price,freight_value,product_category_effect,customer_loc_effect,customer_state_effect,seller_id_effect,distance_between_two_points,mean_delay_in_handover,volume,number_of_hours,product_weight_g),use="pairwise.complete.obs") %>% knitr::kable()
```

@@

```{r rf model, echo = FALSE}
set.seed(seed=1,sample.kind = 'Rounding')
tuningControl=trainControl(method = "cv", number = 2, p = .8)
rf_fit <- train(number_of_hours~., train_subset, method='rf',tuneGrid = data.frame(mtry = c(1,2,3,4,5,10)),trControl=tuningControl,ntree=250)
#rf_fit <- train(number_of_hours~., train_subset %>% select(-seller_state), method='rf',tuneGrid = data.frame(mtry = seq(5, 50, 10)),trControl=tuningControl,ntree=250) # nodesize = 50,maxnodes=2000,ntree=200

@@

RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
# 173.9907 for mtry 2. Technically there are total of 10 predictors. The seller_state was not of much help. it was actually increasing RMSE. So have ignored it.

@@

RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
# 183.4923 uses GCV.Cp method and select=TRUE

@@

```{r Prepare product_category effect based on target encoding, include = FALSE}
# prepare product_category_effect based on regularisation weight we have identified
prepare_product_category_effect <- function(train_df, weight_to_use) {
  global_mean <- mean(train_df$number_of_hours)

  product_category_effects_list <- lapply(1:10, function(index) {
      train_weight_set <- train_df[-test_weight_indices[,index],]

      product_category_effects <- train_weight_set %>% 
        group_by(product_category_name) %>%
        summarise(mean=mean(number_of_hours),n=n()) %>%
        mutate(mean=(n*mean+weight_to_use*global_mean)/(n+weight_to_use)) %>% 
        select(-n) %>% rename()

      product_category_effects
    }
  )

  pce <- product_category_effects_list[[1]]
  for (index in 2:10) {
     pce <- pce %>% full_join(product_category_effects_list[[index]],by='product_category_name', suffix=c("1",index))
  }

  # wherever value is NA, we have used global mean
  pce <- pce %>% mutate(across(-product_category_name,~if_else(is.na(.x),global_mean,.x)))

  # as different partitions may not have exactly same data set combined the results
  product_category_effects <- tibble(pce[,1], product_category_effect=rowMeans(pce[,-1],na.rm = TRUE))

  #product_category effects are ready
  product_category_effects
}
```

@@

```{r Enhance training and test dataset for product_category effect based on target encoding, include = FALSE}
enhance_datasets_using_product_category_effect <- function(product_category_effects,train_df,test_df,cleanup) {
  #we will use seller id effects computed using training data in test data.
  train_df <- train_df %>% left_join(product_category_effects,by='product_category_name') %>% mutate(product_category_effect=if_else(is.na(product_category_effect),global_mean,product_category_effect))

  test_df <- test_df %>% left_join(product_category_effects,by='product_category_name')

  # cleanup
  if (cleanup) {
    train_df <- train_df %>% select(-product_category_name)
    test_df <- test_df %>% select(-product_category_name)
  }

@@

  # cleanup
  if (cleanup) {
    train_df <- train_df %>% select(-seller_id)
    test_df <- test_df %>% select(-seller_id)
  }

@@

  print('Enhancing data for product category effects')
  weight_to_use <- target_encoding_weights %>% 
    filter(effect=='product_category') %>% 
    pull(weight)
  product_category_effects <- prepare_product_category_effect(train_ds,weight_to_use)
  dataset_list <- enhance_datasets_using_product_category_effect(product_category_effects,train_ds,test_ds,cleanup)
  train_ds <- dataset_list[[1]]
  test_ds <- dataset_list[[2]]

@@

# cleanup
training_dataset <- training_dataset %>%
  select(-handed_over_to_carrier_date,-shipping_limit_date,-product_category_name,-seller_id,-customer_state,-customer_loc,-customer_zip_code_prefix,-order_purchase_timestamp,-payment_approved_at)
validation_dataset <- validation_dataset %>%
  select(-handed_over_to_carrier_date,-shipping_limit_date,-product_category_name,-seller_id,-customer_state,-customer_loc,-customer_zip_code_prefix,-order_purchase_timestamp,-payment_approved_at)
```

@@

RMSE((validation_dataset %>% pull(number_of_hours) ),predictions)
# 189.2705 for mtry=2, ntree=500
```

```{r Final model analysis, echo=FALSE}
validation_dataset %>% select(number_of_hours) %>% mutate(diff=(number_of_hours-predictions)) %>% ggplot(aes(diff)) + geom_histogram(binwidth = 1) + scale_x_log10()

validation_dataset %>% select(number_of_hours) %>% mutate(diff=number_of_hours-predictions) %>% summary()

```

```{r Final knn model traing with data scaled, echo=FALSE}
set.seed(seed=1,sample.kind = 'Rounding')
preProcessScalingValues <- preProcess(training_dataset %>% select(-number_of_hours), method = c("center", "scale"))

scaled_train_subset <- predict(preProcessScalingValues, training_dataset) 
scaled_test_subset <- predict(preProcessScalingValues, validation_dataset)

final_knn_fit <- knnreg(number_of_hours~., scaled_train_subset,k = knn_fit$finalModel$k)

final_knn_fit

predictions <- predict(final_knn_fit,scaled_test_subset %>% select(-number_of_hours))

RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
# 203.0848

save(final_knn_fit,file='rda/final_knn_fit.rda')
```
